Determinism, Complexity, and Predictability in Computer Performance

### 2014-02-26 results.tex

 \section{Validation of Primary Findings} % (fold)
 \label{sec:results}

@@

In this section we validate the two key primary findings of this work which we introduced in Section \ref{sec:intro}:

@@

\item The existence of predictable structure in noisy real-valued time series is quantifiable by WPE and as a result WPE is correlated with prediction accuracy (MASE) of an ideal predictor.

%\subitem [[Simply reminders not to be included]]WPE can quantify when a noisy real-valued time series is predictable.[[I am unsatisfied with predictable in this sentence, need a better word to say "better than random walk" or ``able to be forecast effectively" or ``has the structural capacity to transmit information in a way that the time series can be effectively forecast"

%\item The existence of complex structure in a noisy real-valued time series is quantifiable and this type of complexity is directly correlated with predictability

\item The way structure/information/complexity is processed internally by a given process plays a crucial role in predictability.

%\subitem [[Simply reminders not to be included]]We will have shown that the existence whether linear or nonlinear is picked up on with WPE but this point gets at whether the prediction model can use the structure or not (linear can't use nonlinear structure). The right to left shifts in \col and some of the \svd regimes and the lack of shift in \gcc illustrate this nicely.

%\subitem [[Simply reminders not to be included]]This is the linear vs nonliear vs random we see with the right to left shifts with lower complexity time series.

@@

%\item Quantifying structured and unstructured complexity is nontrivial in the case of real-valued noisy time series but WPE does this. [[talk about again here but justify in information theory section]]

As we discussed in Section \ref{sec:meaComplex} distinguishing structured from unstructured complexity in the case of real-valued time series is non trivial, i.e., distinguishing when a complex signal omits predictable structure and when the signal is effectively random is not a trivial task. As described in Section \ref{sec:intro}, for a practitioner this can be frustrating because it can be nearly impossible to find the source of faulty predictions: Is it simply that we need to use a more ideal (possibly more advanced) predictor or is it the case that the time series is simply so complex that using a simple (yet inconsistent) forecast strategy such as random walk is the best we can do. Fortunately, weighted permutation entropy (WPE) allows us to make this distinction for noisy real-valued time series.

In Figure~\ref{fig:pred_vs_wpe} we plot the best prediction (i.e., the lowest MASE over all 3 methods\footnote{LMA,ARIMA,na\"ive} over all 15 runs of that program) for each of the 8 programs under investigation. 
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.75\textwidth]{figs/prediction_vs_entropy}
  \caption{The best MASE among all runs and prediction methods vs weighted
  permutation entropy. For each of these, the word length used is $6$. The
  dashed line is a least-squares linear fit of all the points except for {\tt
  SVD$_1$} which we have excluded for reasons explained in the text.}
  \label{fig:pred_vs_wpe}
\end{figure}
What we see here is that the relationship between prediction accuracy (MASE) and the weighted permutation entropy is much as we conjectured: The existence of predictable structure in noisy real-valued time series as quantified by low to moderate WPE  is correlated with prediction accuracy (MASE) of the ``best"predictor\footnote{We do not claim to have found the ideal predictor for each signal, simply the best predictor over what we used; Which we believe to be a fair sampling of standard prediction techniques for this type of signal.}. We will not further validate this claim through some examples.

To further validate this finding we present an in-depth analysis of three examples (\col, \svdfive and \gcc ) which nicely cover the range of complexity from low to high respectively. In Figures~\ref{fig:lma_vs_arima} and \ref{fig:arima_pred_vs_ent} we plot the MASE of LMA and ARIMA (respectively) against the weighted permutation entropy of each of the 15 runs of all 8 programs. We also list all of these values in Table \ref{tab:error}

\begin{table}[htdp]
\caption{MASE distributions for 1-step predictions at a 10\% prediction horizon over 15 runs for each signal and average wpe at word length 6 for each signal.}
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
                   & MASE LMA    & MASE ARIMA &MASE na\"{i}ve   & $l=6$ \\
\hline
\gcc                  & $ 1.5296\pm 0.0214$ & $1.8366 \pm0.0157 $ & $0.9510 \pm 0.0011$ & $0.9430 \pm 0.0013$ \\

\col           & $ 0.0500 \pm0.0018  $ & $0.5989  \pm 0.2114 $ & $0.5707\pm0.0017$&  $0.5131 \pm 0.0034$ \\

\svd Reg. 1     & $ 0.8273\pm 0.0755$ & $ 0.7141\pm 0.0745 $ & $2.6763\pm4.3282$&  $0.9572 \pm 0.0156$ \\
\svd Reg. 2     & $1.2789 \pm0.0196 $ & $2.1626 \pm0.0265 $ &  $3.0543\pm0.0404$ &   $0.8464 \pm0.0044$ \\
\svd Reg. 3       & $0.6192 \pm0.0209 $ & $0.7129 \pm 0.0096 $ & $31.3857\pm 0.2820$ &  $0.7157 \pm 0.0056$ \\
\svd Reg. 4     & $ 0.7789\pm0.0358 $ & $0.9787 \pm0.0321 $ & $2.6613\pm0.0739$ & $0.8246 \pm 0.0077$ \\
\svd Reg. 5     & $ 0.7177\pm 0.0483 $ & $2.3700  \pm 0.0505 $ & $20.8703 \pm 0.1915$&  $0.6776 \pm 0.0068$ \\
\svd Reg. 6     & $ 0.7393\pm 0.0682 $ & $ 1.4379\pm 0.0609$ & $2.1967\pm0.0830$&  $0.7475 \pm 0.0106$ \\
\hline
\end{tabular}
\end{center}
\label{tab:error}
\end{table}%

\begin{figure}[htbp]
  \centering

  \begin{subfigure}{0.5\textwidth}
    \includegraphics[width=\textwidth]{figs/LMA_prediction_vs_entropy}
    \caption{The MASE of LMA vs weighted permutation entropy. }
    \label{fig:lma_pred_vs_ent}
  \end{subfigure}%
  \begin{subfigure}{0.5\textwidth}
    \includegraphics[width=\textwidth]{figs/ARIMA_prediction_vs_entropy}
    \caption{The MASE of ARIMA vs weighted permutation entropy. }
    \label{fig:arima_pred_vs_ent}
  \end{subfigure}
\caption{
For each of these, the word length used is $6$. The MASE values for LMA against ARIMA. The dashed line is the identity, delineating the traces for which either LMA or ARIMA performed better. All traces except those from \svd$_1$ lie above the line, indicating that LMA is better suited prediction method for the computer traces considered.}\label{fig:lma_vs_arima}  
\end{figure}

We will start with the lowest complexity program: \col. Conceptually this program is very simple but due to computer design choices this program can result in highly complex even chaotic time series traces, see \cite{mytkowicz09}. If we forecast this signal using an out-of-the-box linear forecaster (ARIMA) or an out-of-the-box constant predictor we receive MASE ($0.59 \pm 0.21$ and $0.57 \pm$ 0.001 respectively) which suggest that on average we can only predict this signal twice as well as the simple random walk prediction. This would suggest that there is some predictive structure present but not that much. However, if we quantify the predictive structure present in this signal using WPE we see that this signal is highly structured with a WPE of ($0.51 \pm 0.003$.) The existence of structured which we quantified with WPE implies that an ideal predictor capable of using this signals information should be able to predict this signal much better than the random walk due to the high amount of predictive structure or equivalently the low complexity. This is in fact the case. \col is a complex signal, i.e., chaotic, but it is complex in a structured way. Using a prediction technique that can process nonlinear information yields forecasts with a MASE of $0.05 \pm 0.001$, this is a prediction 20 times more accurate than a random walk forecast. The existence of predictive structure as quantified by WPE could be precisely the information needed to continue searching for an ideal predictor like LMA.

Maybe even more convincing than a low complexity signal like \col is a moderately complex signal like \svdfive. This signal omits a WPE of $0.677 \pm 0.006$, this is a significantly higher complexity than \col but still implies the existence of a great deal of predictive structure. However, this increase in complexity is enough to really throw off the constant and linear prediction strategies. The MASE scores for ARIMA and na\"ive imply 2.5 and 20 times worse prediction than using a simple random walk forecast. This may be discouraging to someone using an out-of-the-box predictor that the signal is too complex to predict. However by \emph{knowing} that there exists predictable  in this signal we can try to use a more advanced technique, in particular LMA. If we do this we get forecasts which score 1.5 times better than the random walk. Showing that this usable structure (although hidden to the linear and constant method) was quantifiable by WPE.

Finally, we consider the most complex program: \gcc. The complexity of \gcc is $0.94 \pm 0.001$. This implies that almost no predictive structure exists in this structure. Recall that a WPE $\approx 1$ are similar in complexity to white noise or a random walk process as they do not transmit any information from past to future. This would mean that the WPE of $>0.95$ would suggest little to no predictive structure exists and that the signal is simply too complex. In this case it should be safe to assume that a random walk forecast would be (on-average) superior to forecast methods which required structure or information to intelligently predict the next time step. This is in fact what we see, with each prediction scheme we used we see a MASE which implies (on-average) 1.5-1.8 times worse predictions that performing a simple random walk forecast.

There is a single outlier in this study which is \svdone. This signal has a WPE of $0.95 \pm 0.02$ and both LMA and ARIMA do slightly better than random walk ($0.82 \pm 0.08$ and $0.7141 \pm 0.07$ respectively). The na\"ive method did significantly worse than random walk and was highly inconsistent with an average MASE of 2.67 but a standard deviation of 4.33 due to the massive right tail this distribution exhibited in reaction to wildly inconsistent forecasts. We do not believe this outlier is really a problem but it does bring about a potential weaknesses of this method. MASE is scaled by a random walk forecast which can be very inconsistent in the case of drastic signal changes which occur on a regular basis, e.g., a signal which oscillates from the top to the bottom of the signals scale at every time click. In this case a random walk forecast would have error with magnitude of the scale of the signal at every forecast, since guessing the last step will be exactly out of phase with the new forecast. A signal like this would normally cause a very low WPE much like \col and a very high Random Walk error resulting in low MASE for methods that can take into account this structure like ARIMA or LMA.

To understand why this causes \svdone to be an outlier we examine Figure \ref{fig:svdone-ts} in more detail. This is a small snippet of the time series of \svdone to illustrate the strucutre of this signal. We provide a small snippet as opposed to the full time series simply for visual clarity.
\begin{figure}[htbp]
  \centering
    \includegraphics[width=\textwidth]{figs/svdonets2}
\caption{ For each of these, the word length used is $6$. The MASE values for LMA against ARIMA. The dashed line is the identity, delineating the traces for which either LMA or ARIMA performed better. All traces except those from \svd$_1$ lie above the line, indicating that LMA is better suited prediction method for the computer traces considered.}\label{fig:svdone-ts}  
\end{figure} 
As can be seen in Figure \ref{fig:svdone-ts}, \svdone has a highly unstructured small band on top, which is approximately 0.05 instructions per cycle wide. Approximately 80\% of the signal exists within this band of noise, this causes the WPE to be driven up as many words are being omitted by this band of the signal. The remaining 20\% of the signal are drops in instructions per cycle away from this band, ranging in size from 0.1 to 0.6 instructions per cycle\footnote{Some of the large drops are not shown for the sake of clarity of the smaller scale}. These drastic drops seem to happen every 5-7 points but the drop locations are inconsistent as well as the drops magnitude (ranging from 0.1-0.6 instructions per cycle). As discussed in Section \ref{sec:RWMethod} drops of this fashion are the bane of random walk forecasting. At every one of these drops a random walk predictor would produce two forecasts with an error 2-12 times the magnitude of the smaller band, i.e., about 40\% of the time the random walk forecast would predict with error 2-12 times the magnitude of the scale exhibited by $\approx$80\% of the signal. This inconsistency in random walk would cause cause random walk error to be high and as a result MASE scores to be artificially low. The combination of a band of noise with a high number of random drops is a weakness of this method which is present in \svdone and as a result makes this signal an outlier.

[[Joshua: I am not sure if this paragraph is strong enough or even worth having]]These results support that time series with low to moderate complexity ($0\le WPE \le 0.85$) can be predicted more efficiently than a na\"ive random walk \emph{and} that complexity can be qualitatively measured for a real-valued noisy time series using WPE. This will allow practitioners to stop spinning their wheels in the case of signals who are simply better predicted with a simple strategy like random walk. The analysis of these results illuminates an interesting point: The way structure, information and complexity are processed by a generating process plays a crucial role in the success of a given prediction scheme.

@@

Other noteworthy features of the LMA and ARIMA results are the cluster locations
and distributions. The WPE values of each run of any particular program tend to
have little variance, leading to the clusters in
Figures~\ref{fig:arima_pred_vs_ent} and \ref{fig:lma_pred_vs_ent} to be fairly
constrained in the $y$ direction. For most traces, the LMA and ARIMA variance is
low as well, resulting is small, tight clusters. The ARIMA MASE values of the \col
traces, however, have a large variance resulting in the spread seen in
Figure~\ref{fig:arima_pred_vs_ent}. Not only are the MASE values of that cluster
bad, in that other predictors vastly outperform it, but they are inconsistent.
Furthermore, since LMA can predict nonlinear behavior while ARIMA cannot, we
see that the clusters in Figure ~\ref{fig:arima_pred_vs_ent} are mostly further to
the right than those in Figure~\ref{fig:lma_pred_vs_ent}.

>>>>

 \section{Predictability, Complexity, and Permutation Entropy} % (fold)
 \label{sec:results}
% I changed the title of this section because there are lots of
% results in the modeling section too.  The results HERE are about the
% relationship between predictability and WPE

@@

In this section, we offer an empirical validation of the two key
conjectures introduced in Section \ref{sec:intro}, namely:

@@

\item that the weighted permutation entropy (WPE) of a noisy
  real-valued time series is correlated with prediction
  accuracy---i.e., that the predictable structure in a time-series
  data set can be quantified by its WPE.

%\subitem [[Simply reminders not to be included]]WPE can quantify when
%a noisy real-valued time series is predictable.[[I am unsatisfied
%with predictable in this sentence, need a better word to say "better
%than random walk" or ``able to be forecast effectively" or ``has the
%structural capacity to transmit information in a way that the time
%series can be effectively forecast"

%\item The existence of complex structure in a noisy real-valued time
%series is quantifiable and this type of complexity is directly
%correlated with predictability

\item that the way information is processed internally by a system is
  correlated with the effectiveness of a given predictor on
  time-series data from that system

%\subitem [[Simply reminders not to be included]]We will have shown
%that the existence whether linear or nonlinear is picked up on with
%WPE but this point gets at whether the prediction model can use the
%structure or not (linear can't use nonlinear structure). The right to
%left shifts in \col and some of the \svd regimes and the lack of
%shift in \gcc illustrate this nicely.

%\subitem [[Simply reminders not to be included]]This is the linear vs
%nonliear vs random we see with the right to left shifts with lower
%complexity time series.

@@

%\item Quantifying structured and unstructured complexity is
%nontrivial in the case of real-valued noisy time series but WPE does
%this. [[talk about again here but justify in information theory
%section]]

The experiments below involve the three different prediction methods
described in Section~\ref{sec:compModel}---na\"ive, ARIMA, and
LMA---applied to time-series data from eight different systems: {\tt
  col\_major}, {\tt 403.gcc}, and the six different regimes of the
{\tt dgesdd} signal in Figure~\ref{fig:wwpe}.  The objective of these
experiments was to explore how prediction accuracy is related to WPE,
and how that relationship depends on the generating process and the
prediction method.  Working from the first 90\% of each signal, we
generated a prediction of the last 10\% of that signal using each of
the methods, as described in Section~\ref{sec:accuracy}, then
calculated the MASE value of those predictions.  We also calculated
the WPE of each time series, as described in
Section~\ref{sec:meaComplex}, using a wordlength of six.  In order to
assess the run-to-run variability of these results, we repeated all of
these calculations on 15 separate trials: i.e., 15 different runs of
each program.  Figure~\ref{fig:wpe_vs_mase_best} shows the {\sl best}
of these 45 predictions for each system: i.e., the lowest error over
all 15 trials and all three methods for each program.  The WPE is
plotted against the corresponding MASE value in order to bring out the
correlation between these two quantities.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/prediction_vs_entropy}
  \caption{Weighted permutation entropy vs. mean absolute scaled error
    (MASE) of the best prediction of each system.
% 
% (across all trials and all prediction methods) 
% 
The dashed line is a least-squares linear fit of all the points except
{\tt dgesdd$_1$}, which we have excluded for reasons explained in the
text.  [[Ryan, is that correct?]]}
  \label{fig:wpe_vs_mase_best}
\end{figure}

The best-case prediction error for each of these eight systems is
roughly proportional to the weighted permutation entropy, which is
consistent with our first conjecture.  The dashed line in the figure,
a linear least-squares fit to these best-case errors---with the
exception of {\tt dgesdd$_5$}, for reasons described below---captures
this rough proportionality.  This is not a formal result.  The three
methods used here were chosen to span the space of standard prediction
strategies, but they do not, of course, cover that space exhaustively.
Our goal here is an empirical assessment of predictability and
complexity, not formal results about a ``best'' predictor for a given
time series.  There may be other methods that produce lower MASE
values than those in Figure~\ref{fig:wpe_vs_mase_best}, but the
sparseness of the points in the upper-left and lower-right quadrants
of this plot strongly suggests that the underlying predictability of a
time series is inversely correlated with its WPE.  The rest of this
section describes our results in more detail, including the measures
taken to assure meaningful comparisons across methods, trials, and
programs, and elaborates on the meaning of the dashed line in the
figure.

Figure~\ref{fig:wpe_vs_mase_all} shows WPE vs. MASE plots for all 360
experiments.
\begin{figure}
  \centering
    \includegraphics[width=0.6\textwidth]{figs/ARIMA_prediction_vs_entropy}
    \includegraphics[width=0.6\textwidth]{figs/ARIMA_prediction_vs_entropy}
    \includegraphics[width=0.6\textwidth]{figs/LMA_prediction_vs_entropy}
\caption{WPE vs. MASE for all trials, methods, and systems.  \svdone,
  \svdthree, and \svdfive are omitted from the top plot for scale
  reasons; their MASE scores are $2.6763\pm4.3282$, $31.3857\pm
  0.2820$, and $20.8703 \pm 0.1915$ respectively.  {\color{red}Replace
    the top plot with a plot of the naive results---but running from 0
    to 3.1 on the horizontal axis and excluding SVD 1, 3, and 5}.  The
  dashed line is the same as in the previous figure [[Ryan, is this
      correct?]].  Numerical values, including means and standard
  deviations of the errors, can be found in Table~\ref{tab:error} in
  the appendix.
% 
%  [[All traces except those from \svd$_1$ lie above the line,
%      indicating that LMA is better suited prediction method for the
%      computer traces considered.]]  [[This comes {\bf long} before the
%      associated discussion in the text, and it's not what this figure
%      is trying to show here.]]
% 
}
    \label{fig:wpe_vs_mase_all}
\end{figure} 
There are 15 points in each cluster, one for each trial\footnote{A
  trial involves collecting a time series from the corresponding
  system, predicting the last 10\% of that time series, and
  calculating the MASE score.}.  (The points in
Figure~\ref{fig:wpe_vs_mase_best} are the leftmost points in any of
the three clusters for the corresponding system.)  Most of the
clusters are quite tight.  Some are flat and wide because the WPE
values of each run of any particular program tend to have little
variance; for most traces, the LMA and ARIMA variance is low as well,
resulting in small, tight clusters.  The {\sl spatial distribution} of
the clusters is quite different in the three plots: LMA is tight
around the dashed line, whereas naive and ARIMA are [[need to fill in
    words here when get the third image]].  This is a geometric
validation of the second conjecture in this paper, as described in the
following paragraphs.

Though \col is a very simple program, its dynamics are actually quite
complicated, as discussed in Section~\ref{sec:intro}.  Recall from the
right-hand column of Figure~\ref{fig:gcc_vs_col} that the naive and
ARIMA prediction methods perform poorly on this signal (MASE values
$0.57 \pm$ 0.001 and $0.59 \pm 0.21$, respectively).  That is, the
errors produced by these methods are a little over half as large as
the errors produced by the random-walk method, which simply uses the
current value as the prediction.  However, the WPE value for the \col
trials is $0.51 \pm 0.003$: i.e., right in the middle of the
complexity spectrum described in Section~\ref{sec:intro}, implying
that roughly half of the signal is predictive structure.  This
disparity---WPE values that suggest a high rate of forward information
transfer in the signal, but predictions with low MASE
scores---manifests geometrically in the top two images in
Figure~\ref{fig:wpe_vs_mase_all}, where the \col clusters are far to
the right of the dashed line.  This is important; it says that the
methods aren't leveraging that information very well.  This is not
surprising, really.  The dynamics of \col may be complicated, but they
are not unstructured.  This signal is nonlinear, deterministic, and
chaotic.  Indeed, if you use a prediction technique that uses a
nonlinear model (LMA) you get a MASE of $0.05 \pm 0.001$---i.e., 20
times more accurate than a random-walk forecast.  Note the position of
the \col cluster in the bottom image in
Figure~\ref{fig:wpe_vs_mase_all}: much closer to the dashed line.
From a practitioner's standpoint, what this means is that being below
that line tells you that there really IS usable structure in the time
series, that you could do better, and that you should use a better
method.  Note, too, that the \col points are clustered in LMA but
spread out horizontally in ARIMA, far to the right of the dashed line
(i.e., much worse than the other methods).  We need to come up with a
hypothesis about this spread and skew, even if it's totally tentative.

The WPE of \svdfive ($0.677 \pm 0.006$) is higher than that of {\tt
  col\_major}.  This indicates that the rate of forward information
transfer of the underlying process is lower, but that time-series data
from this system still contain a significant amount of structure that
can, in theory, be leveraged to predict the future course of the time
series.  However, the effectiveness of any prediction strategy will
depend on how well it leverages the available information.
% 
% (This is the basis for the second conjecture above.)  
% 
This issue becomes apparent in the case of \svdfive, as is clear from
the positions of the corresponding clusters in
Figure~\ref{fig:wpe_vs_mase_all}.  The naive and ARIMA prediction
strategies in the top two plots produce MASE scores of $2.3700 \pm
0.0505 $ and $20.8703 \pm 0.1915$, respectively: that is, 2.37 and
20.87 times worse than a simple random walk forecast of the same
signals.  [[This is big because the amount of variance in this signal
    makes it a bad idea to just guess the mean.  The same thing is
    true of \svdthree.]] [[Joshua: can you say that better---in a way
    that is true of both of these, and distinguishes them from all of
    the signals that have lower values?]] Again, the positions of the
ARIMA and LMA clusters---significantly below and to the right of the
dashed line---should suggest to a practitioner that there is more
structure in this signal than the method is able to leverage, so s/he
can do better and should try a different prediction method.  In this
case, the LMA method produces a MASE score of $ 0.7177\pm 0.0483 $ and
a cluster of results on the WPE-MASE plot that is much closer to the
dashed line.  This validates our second conjecture: the naive and
ARIMA prediction methods, which use constant and linear models,
respectively, cannot capture or reproduce the way in which the
\svdfive system processes information, but the nonlinear LMA method
can.

The WPE of \gcc is much higher: $0.94 \pm 0.001$.  This system
transmits very little information forward in time and provides almost
no structure for prediction methods to work with.  What we see: naive
does the best, then LMA, then ARIMA.  Hypothesis: that that 5\% of the
structure is nonlinear, so LMA can use it better than ARIMA can.
Naive does better than LMA because it's filtering out the noise.

\svdone behaves very differently than the other seven systems in this
study: though its WPE is very high ($0.95 \pm 0.02$), two of the three
prediction methods (LMA and ARIMA) do quite well: $0.82 \pm 0.08$ and
$0.7141 \pm 0.07$ MASE scores, respectively.  Accordingly, the
clusters of \svdone points in the bottom two images in
Figure~\ref{fig:wpe_vs_mase_all} are well above the trend followed by
the other seven clusters.  Also, the MASE scores of the predictions
produced by the na\"ive method are highly inconsistent: $2.67 \pm
4.33$.  These effects, we believe, are artifacts of the way MASE is
calculated.  Recall that MASE scores are scaled relative to a random
walk forecast, which uses the current value as the prediction.  This
strategy works very badly on signals with frequent, large, rapid
transitions.  Consider a signal that oscillates from one end of its
range to the other at every step.  A signal like this will have a low
WPE, much like \col.  However, a random-walk forecast of this signal
will be 180 degrees out of phase with the true continuation.  Since
random-walk error appears in the denominator of the MASE score, this
effect can shift points leftwards on a WPE vs. MASE plot, and that is
exactly why the \svdone clusters for ARIMA and LMA are above the
dashed line in Figure~\ref{fig:wpe_vs_mase_all}.  The \svdone time
series, which is shown in Figure~\ref{fig:svdone-ts},
\begin{figure}[htbp]
  \centering
    \includegraphics[width=\textwidth]{figs/svdonets2}
\caption{A small portion of the \svdone time series}\label{fig:svdone-ts}
\end{figure} 
is not quite to the level of the worst-case signal described above,
but it poses a serious challenge to random-walk prediction.  It is
dominated by a white-noise regime (between 1.86 and 1.88 on the
vertical scale in the Figure), punctuated by short excursions above
1.9.  In the former regime, which makes up more than 80\% of the
signal, there are frequent dips to 1.82 and occasional larger dips
below 1.8.  These single-point dips are the bane of random walk
forecasting.  In this particular case, roughly 40\% of the forecasted
points are off by the width of the associated dip, which skews the
associated MASE scores.  Signals like this are also problematic for
the naive prediction strategy, since the outliers have significant
pull on the mean, which compounds the effect of the skew in the
scaling factor and creates a large spread in the \svdone MASE values.
For all of these reasons, we view \svdone as an outlier and thus
exclude it from the fit calculation of the dashed line in
Figure~\ref{fig:wpe_vs_mase_best}.

@@

The big picture: Furthermore, since LMA can predict nonlinear behavior
while ARIMA cannot, we see that the clusters in Figure
~\ref{fig:arima_pred_vs_ent} are mostly further to the right than
those in Figure~\ref{fig:lma_pred_vs_ent}.
